Fixing Vulnerabilities Hinders Maintainability

### 2019-02-04 paper/ICPC19/icpc19.tex
more fixes

\lstdefinestyle{PHPStyle} {
    language=PHP,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

@@

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{mygreen}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
 <ul>
   <li class='error'>
      <?php echo $l->t('Cloud not found');?><br/>
-     <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+     <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?><?php 
   </li>
 </ul>

\end{lstlisting} \Luis{no listing lst:fix o que achas de fazer linebreak para
cada uma das tags? assim nao havi o problema de ter a mesma linha dividida e ficava mais facil de ler:\\
\texttt{<p class='hint'>}\\
\texttt{~~<?php }\\
\texttt{</p>}\\
Also: reduz numero de espacos para cada tab -- 2 é suficiente.
}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}\Luis{como usamos numeros na enumeracao que descreve a
figura, esses numeros tb deveriam estar na figura.}:

@@

    \item \textbf{Miscellaneous (195 cases)} This pattern comprises several
    other categories of security refactorings (e.g., \textit{path traversal and
    buffer overflow})\Luis{exemplos ja nao precisa estar aqui porque sao
    enumerados logo a seguir}. It contains several other security refactorings
    that do not have their own pattern yet. The patterns not satisfying the
    size requirement of more than $20$ refactorings are the following ones:
    \textit{Resource Leak} (19), \textit{SHA-1 Hash Function} (1),
    \textit{Broken Access Control} (2), \textit{Sensitive Data Exposure} (18),
    \textit{Insufficient Attack Protection} (4), \textit{Overflow} (18),
    \textit{Path Traversal} (19).\Luis{Acho que este item pode ser reescrito de forma mais clara. sugestao:
    This category comprises the security refactorings that (1) do not relate to
    any particular pattern or (2) relate to patterns that do not occur in at
    least 20 refactorings: \textit{Resource Leak} (19), \dots} \end{itemize}

@@

The dataset contains security flaws for more than $13$ different languages,
being PHP ($39\%$) and C ($24\%$) the most prevalent ones (see Figure~\ref{fig:lang}).
To classify the applications in the dataset, we use a taxonomy for open-source
software from previous work~\cite{7816479}. Figure~\ref{fig:domain}
presents the projects domain distribution: \textit{Application Software} ($25$),
software that provides end-users with functional systems; \textit{Web libraries
and frameworks} ($20$); \textit{System Software} ($18$), software that provides
services and infrastructures (e.g., operating systems, servers and databases);
\textit{Software Tool} ($13$), software that supports development (e.g.,
programming languages, compilers, package managers, IDEs); and, \textit{Non-web
libraries and frameworks} ($18$), for desktop and mobile software.
%
Table~\ref{tab:dataset} presents the descriptive statistics of the $94$
open-source projects involved in this study, including the number of forks, the
number of stars, the number of watchers, the number of contributors, the number
of commits, the number of branches, the number of releases,  size (cf. GitHub, it
is the size of the whole repository, including all of its history, in
kilobytes), the number of issues, and the number of pull requests.

@@

One of the guidelines with which the project does not comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}): \emph{Write Simple Units of Code}. SIG defines \emph{Units} as the smallest groups of code that can be maintained and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and constructors in Java). BCH evaluates this guideline based on the number 
\Luis{this seems unfinished. i think we ve already mention this Mccabe metric somewhere else.}

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test \Rui{again} per pattern indicated statistical
evidence supporting that the maintainability of open-source software decreases
for $4$ different patterns: \emph{Broken Autentication} (p-value = $0.018$),
\emph{Denial-of-Service}\Luis{em varios pontos deste paper este pattern era
referido com a sigla DoS. Por uma questao de consistencia com os outros
patterns sugiro que se escreva sempre por extenso (ou entao sempre com a
sigla).} (p-value = $0.029$), \emph{Memory Leak} (p-value = $0.009$) and
\emph{Miscellaneous} (p-value = $3.07$x$10^{-4}$). With statistical
significance, results demonstrate evidence that maintainability is not
negatively impacted when removing \emph{Components with Known Vulnerabilities}
(p-value = $7.96$x$10^{-4}$) and fixing \emph{Cross-Site Scripting
Vulnerabilities} (p-value = $0.001$). \emph{Injection} and \emph{Cross-Site
Request Forgery} produced more cases where the refactorings had a negative
impact in the maintainability. However, the results obtained for these patterns
are not statistically significant.

>>>>

\lstdefinestyle{PHPStyle} {
    language=PHP,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    stepnumber=1,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

@@

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{green}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
 <ul>
  <li class='error'>
   <?php echo $l->t('Cloud not found');?>
   <br/>
   <p class='hint'>
-    <?php if(isset($_['file'])) echo $_['file']?>
   </p>
   <p class='hint'>
+    <?php if(isset($_['file'])) echo htmlentities($_['file'])?>
   </p> 
  </li>
 </ul>

\end{lstlisting} % \Luis{no listing lst:fix o que achas de fazer linebreak para
% cada uma das tags? assim nao havi o problema de ter a mesma linha dividida e ficava mais facil de ler:\\
% \texttt{<p class='hint'>}\\
% \texttt{~~<?php }\\
% \texttt{</p>}\\
% Also: reduz numero de espacos para cada tab -- 2 é suficiente.
% }
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology is comprised of the following steps, as illustrated in
Figure~\ref{fig:met}:

@@

    \item \textbf{Miscellaneous (195 cases)} This pattern comprises several other categories
    of security refactorings (e.g., \textit{path traversal and buffer
    overflow}). It contains several other security refactorings that do not have
    their own pattern yet. The patterns not satisfying the size requirement of
    more than $20$ refactorings are the following ones: \textit{Resource Leak} (19),
    \textit{SHA-1 Hash Function} (1), \textit{Broken Access Control} (2),
    \textit{Sensitive Data Exposure} (18), \textit{Insufficient Attack Protection} (4),
    \textit{Overflow} (18), \textit{Path Traversal} (19).
\end{itemize}

@@

The dataset contains security flaws for more than $13$ different languages,
being PHP ($39\%$) and C ($24\%$) the most prevalent ones (see Figure~\ref{fig:lang}).
To classify the applications in the dataset, we use a taxonomy for open-source
software from previous work~\cite{7816479}. Figure~\ref{fig:domain}
presents the projects domain distribution: \textit{Aplication Software} ($25$),
software that provides end-users with functional systems; \textit{Web libraries
and frameworks} ($20$); \textit{System Software} ($18$), software that provides
services and infrastructures (e.g., operating systems, servers and databases);
\textit{Software Tool} ($13$), software that supports development (e.g.,
programming languages, compilers, package managers, IDEs); and, \textit{Non-web
libraries and frameworks} ($18$), for desktop and mobile software.
%
Table~\ref{tab:dataset} presents the descriptive statistics of the $94$
open-source projects involved in this study, including the number of forks, the
number of stars, the number of watchers, the number of contributors, the number
of commits, the number of branches, the number of releases,  size (cf. GitHub, it
is the size of the whole repository, including all of its history, in
kilobytes), the number of issues, and the number of pull requests.

@@

One of the guidelines with which the project does not comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}): \emph{Write Simple Units of Code}. SIG defines \emph{Units} as the smallest groups of code that can be maintained and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and constructors in Java). BCH evaluates this guideline based on the number

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test \Rui{again} per pattern indicated statistical evidence
supporting that the maintainability of open-source software decreases for $4$
different patterns: \emph{Broken Autentication} (p-value = $0.018$),
\emph{Denial-of-Service} (p-value = $0.029$), \emph{Memory Leak} (p-value = $0.009$) and
\emph{Miscellaneous} (p-value = $3.07$x$10^{-4}$). With statistical significance,
results demonstrate evidence that maintainability is not negatively impacted
when removing \emph{Components with Known Vulnerabilities} (p-value = $7.96$x$10^{-4}$)
and fixing \emph{Cross-Site Scripting Vulnerabilities} (p-value = $0.001$).
\emph{Injection} and \emph{Cross-Site Request Forgery} produced more
cases where the refactorings had a negative impact in the maintainability.
However, the results obtained for these patterns are not statistically
significant.
